Writing Trash Bin

### 2024-03-14 paper.qmd
small edits

A large drive of the field of uncertainty visualisation, has been the idea that "uncertainty visualisation" (defined as one word) is somehow different to "uncertainty" "visualisation" (separately defined as two words). When "uncertainty visualisation" is defined as a single word we lose some nuance in our understanding of the driving force behind some issues, and problems with visualisation and problems with uncertainty become conflated. This appears in the literature as a mountain of seemingly conflicting statements about what is, or is not, an "uncertainty visualisation". For example @Leland2005 mentions that popular graphics, such as pie charts and bar charts omit uncertainty, however at least one or both of these charts are used in most "uncertainty visualisation" experiments [@Ibrekk1987] (*Cite Gap __*). @Wickham2011 suggests their product plot framework, which includes histograms, should have a way to measure uncertainty, but does not consider that a histogram is *already* a depiction of mass and would already be considered an uncertainty visualisation were our statistic of interest the population mean.

These conflicting ideas and lacking definition of what is means for a visualisation to be an "uncertainty visualisation" is a conerstone of some clearly questionable results within this field.

@@

> Edit Notes: Similar to the "similar works" section, I think this section would be better off with a short sentance about each paper and comments on their similarities because a paragraph for each paper is too much

@@

### Uncertainty as signal

@@

## Uncertainty versus Risk
Uncertainty is noise. This may be an obvious statement, but what is not obvious is that the existing literature sees noise as a signal. When we directly ask "What is the varinance of this variable" the variance *becomes* the signal. Noise and signal are not an inherrent property of variance, or meta-data or the aspects of an analysis that are usually secondary considerations, it is a property of the data that *promotes* a message we are trying to infer and the data that *supresses* it. The replication problem in uncertainty visualisation, where experiments are simply repeating established visualisation work, comes from the misunderstandings established in previous sections. Uncertainty is poorly defined, therefore authors do not understand it is not an inherrent property of the data and is actually related to a particular inference question, this incorrect view that uncertainty is always variance or probabilities leads to the class of "uncertainty visualisations" that depict different mathematical objects or visual features with little in the sense of consistent rules, this complicated class has caused visualisation authors to miss that they are simply repeating the established evaluation results from normal information visualisations. This situation is a bit of a mess. It also highlights a unique and facinating problem faced by uncertainty visualisation. If asking direct questions about uncertainty causes us to treat it as a signal, how do we evaluate uncertainty as *noise*?

>>>>

A large drive of the field of uncertainty visualisation, has been the idea that "uncertainty visualisation" (defined as one word) is somehow different to "uncertainty" "visualisation" (separately defined as two words). When "uncertainty visualisation" is defined as a single word we lose some nuance in our understanding of the driving force behind some issues, and problems with visualisation and problems with uncertainty become conflated. This appears in the literature as a mountain of seemingly conflicting statements about what is, or is not, an "uncertainty visualisation". For example @Leland2005 mentions that popular graphics, such as pie charts and bar charts omit uncertainty, however at least one or both of these charts are used in most "uncertainty visualisation" experiments [@Ibrekk1987] (*Cite Gap 10*). @Wickham2011 suggests their product plot framework, which includes histograms, should have a way to measure uncertainty, but does not consider that a histogram is *already* a depiction of mass and would already be considered an uncertainty visualisation were our statistic of interest the population mean.

These conflicting ideas and lacking definition of what is means for a visualisation to be an "uncertainty visualisation" is a cornerstone of some clearly questionable results within this field.

@@

Similar to "uncertainty" exactly what defines an "uncertainty visualisation" is not clear. Often, the visualisaitons are simply categorised according to features, however exactly what *is* or *is not* an uncertainty visualisation is rarely explained.

@@

## Uncertainty as signal

@@

## The current summaries of findings
## Uncertainty versus Risk
Uncertainty is noise. This may be an obvious statement, but what is not obvious is that the existing literature sees noise as a signal. When we directly ask "What is the varinance of this variable" the variance *becomes* the signal. Noise and signal are not an inherrent property of variance, or meta-data or the aspects of an analysis that are usually secondary considerations, it is a property of the data that *promotes* a message we are trying to infer and the data that *supresses* it. The replication problem in uncertainty visualisation, where experiments are simply repeating established visualisation work, comes from the misunderstandings established in previous sections. Uncertainty is poorly defined, therefore authors do not understand it is not an inherrent property of the data and is actually related to a particular inference question, this incorrect view that uncertainty is always variance or probabilities leads to the class of "uncertainty visualisations" that depict different mathematical objects or visual features with little in the sense of consistent rules, this complicated class has caused visualisation authors to miss that they are simply repeating the established evaluation results from normal information visualisations. This situation is a bit of a mess. It also highlights a unique and facinating problem faced by uncertainty visualisation. If asking direct questions about uncertainty causes us to treat it as a signal, how do we evaluate uncertainty as *noise*?

### 2024-03-14 writingdumpster.qmd
small edits

Cite Gap 3: Some consider it to be synonymous with specific terms defined in mathematics, such as probability, variance, error, or precision. Others consider it to be an encompassing umbrella term of which for these mathematically defined objects are only examples, and include concepts that are somewhat related, such as missing values

Cite Gap 3.5: This wide array of uncertainty definitions becomes a snake eating its own tail, as many authors opt to redefine uncertainty themselves[], ignore defining uncertainty entirely [], or pick a definition from a seemingly randomly previously published papers [], all of which make variance in uncertainty definitions even worse

@@

Cite Gap 5: This definition swap happens so subtly that it seems to go unnoticed by authors writing the papers, and the fact that "uncertainty visualisation" methods dont even notice they are exclusively focus on depictions of easily quantifiable uncertainty, such as error [], mass [], or variance []

Cite Gap 6: This has left the expressions of uncertainty that are hard to quantify or visually differentiate untouched, despite many papers calling for their invention

Cite Gap 7: Taxonomies split uncertainty based on an endless stream of arbitrary lines in the sand, such as whether the uncertainty is due to true randomness or a lack of knowledge [@Spiegelhalter2017], if the uncertainty is in the the attribute, spatial elements, or temporal element of the data [], whether the uncertainty is scientific (e.g. error) or human (e.g. disagreement among parties) [], if the uncertainty is random or systematic [], statistical or bounded [], accuracy or precision [], if the uncertainty is about the past or the future [], the stage of the data analysis pipeline it comes from [], the severity of the uncertainty [], how quantifiable the uncertainty is [@Spiegelhalter2017], and many, many, many more examples.

Cite Gap 8: It is commonly noted that each uncertainty must be discussed in isolation, but combining the uncertainty from every stage is near impossible [@Spiegelhalter2017]

@@

>>>>

Cite Gap 3: Some consider it to be synonymous with specific terms defined in mathematics, such as probability [*HERE*], variance[*HERE*], error[*HERE*], or precision[*HERE*]. Others consider it to be an encompassing umbrella term of which for these mathematically defined objects are only examples[*HERE*], and include concepts that are somewhat related, such as missing values[*HERE*]

Cite Gap 3.5: This wide array of uncertainty definitions becomes a snake eating its own tail, as many authors opt to redefine uncertainty themselves[*HERE*], ignore defining uncertainty entirely [*HERE*], or pick a definition from a seemingly randomly previously published papers [*HERE*], all of which make variance in uncertainty definitions even worse

@@

Cite Gap 5: This definition swap happens so subtly that it seems to go unnoticed by authors writing the papers, and the fact that "uncertainty visualisation" methods dont even notice they are exclusively focus on depictions of easily quantifiable uncertainty, such as error [*HERE*], mass [*HERE*], or variance [*HERE*]

Cite Gap 6: This has left the expressions of uncertainty that are hard to quantify or visually differentiate untouched, despite many papers calling for their invention [*HERE*]

Cite Gap 7: Taxonomies split uncertainty based on an endless stream of arbitrary lines in the sand, such as whether the uncertainty is due to true randomness or a lack of knowledge [@Spiegelhalter2017 *HERE*], if the uncertainty is in the the attribute, spatial elements, or temporal element of the data [*HERE*], whether the uncertainty is scientific (e.g. error) or human (e.g. disagreement among parties) [*HERE*], if the uncertainty is random or systematic [*HERE*], statistical or bounded [], accuracy or precision [*HERE*], if the uncertainty is about the past or the future [*HERE*], the stage of the data analysis pipeline it comes from [*HERE*], the severity of the uncertainty [*HERE*], how quantifiable the uncertainty is [@Spiegelhalter2017], and many, many, many more examples.

Cite Gap 8: It is commonly noted that each uncertainty must be discussed in isolation, but combining the uncertainty from every stage is near impossible [@Spiegelhalter2017 *MORE HERE*]

@@

Cite Gap 9: For example @Leland2005 mentions that popular graphics, such as pie charts and bar charts omit uncertainty, however at least one or both of these charts are used in most "uncertainty visualisation" experiments [@Ibrekk1987 *MORE HERE*]

Cite Extra: INCORRECT INTERPRETATION OF STUDIES. People will use the best available information to them. If you ask for a maximum, and give them a confidence interval... they will give you the maximum number that they are aware is still plausible. Instead of being shocked the participants were wrong, it is important to ask "what was required for them to be RIGHT".
  - HAVE SOME PERSPECTIVE. If you are paying people below minimum wage to do your experiment, they don't care. If a problem requires extensive mental labour to solve, they will liekly not do it.
  - This result being reported as "people interpret a confidence interval as a maximum and minimum despite being told it was a confidence interval" is simply not true. It is a false equivalence. 
  - all these studies show is if you give people the wrong information... they will be wrong. Shocking.
